My wonderful paper

### 2024-09-13 _manuscript/_tex/index.tex
explain result universeg

@@

Experienced analysts often have implicit expectation about the outcome
and rely on a few ``directional signs'' to check when the outcome
deviate from those expectation. However, these expectations are rarely
made explicit within the analysis workflow. This makes it challenging
for consumers of the analysis to evaluate the results, since it becomes
difficult to disentangle whether discrepancies arise from differing
expectations or from the use of statistical technique, without running
the analysis themselves. Non-expert analysts, lacking prior knowledge or
instinct, may not have clear expectations of the results. This can lead
to reduced confidence of the analysis and makes it more difficult and
time-consuming to diagnose the cause of the deviation when the results
don't align with expectations.

By explicitly formulating these expectations, an analysis plan can guide
the analysis process, facilitate the communication and comparison with
other analyses and evaluate the validity of the results.

\begin{itemize}
\tightlist
\item
  why it is useful to be explicit about expectation?

  \begin{itemize}
  \tightlist
  \item
    and can be used to guide the analysis process
  \item
    with simulation, the unit tests can be used to construct the
    ``result universe''
  \item
    With formulated expectation, one can generate multiple version of
    the dataset for teaching statistics
  \end{itemize}
\end{itemize}

When the expectation is formulated, we can also construct values fulfill
or not fulfill the expectation. With simulation, this allows to
construct the set of all possible results that can be obtained from the
analysis process, which we called the result universe. With such a
result universe, we can see how the results obtained by the analysts are
situated in all possible results. When accompanied by the result
universe, it is clearer how different factors in the analysis affects
the main results. This can be seen as a direct towards trustworthy data
analysis, where the audience of the analysis to exercise their own
cognitive model \citep{grolemund_cognitive_2014} to evaluate the results
reported.

>>>>

The whole point being how can be build trust in other's analysis. Some
provides scripts to run the analysis but that's not enough. We need to
know what to expect from the analysis and that is often not explicitly
specified.

@@

Experienced analysts often have implicit expectation about the outcome
and rely on a few ``directional signs'' to check when the outcome
deviate from those expectation. However, these expectations are rarely
made explicit within the analysis workflow. This makes it challenging
for consumers of the analysis to evaluate the results, since it becomes
difficult to disentangle whether discrepancies arise from differing
expectations or from the use of statistical technique, without running
the analysis themselves. Non-expert analysts, lacking prior knowledge or
instinct, may not have clear expectations of the results. This can lead
to reduced confidence of the analysis and makes it more difficult and
time-consuming to diagnose the cause of the deviation when the results
don't align with expectations. By explicitly formulating these
expectations, an analysis plan can guide the analysis process,
facilitate the communication and evaluate the validity of the results.

The expectations can be thought of as a set of unit tests used to
validate the results of data analysis. By specifying a range of values
for these tests, multiple versions of the dataset can be generated to
satisfy different sets of plan expectations. This allows us to present
what we called the ``result universe'' -- the complete set of possible
results that can be obtained from one data analysis process. By
visualizing the result universe, data analysis consumers can observe how
changes in expectations affect the results and the range of alternative
outcomes that could arise under different conditions. This enables them
to evaluate the outcomes based on their own plan expectations and gain a
broader perspective on how the actual results produced by analysts fit
within this spectrum of possibilities, promoting transparency and trust
in analysis.

Furthermore, by generating multiple versions of the data, we can emulate
various scenarios within the same context for students to exercise
judgement when conducting data analysis in a classroom setting.

### 2024-09-13 index.qmd
explain result universeg

Experienced analysts often have implicit expectation about the outcome and rely on a few "directional signs" to check when the outcome deviate from those expectation. However, these expectations are rarely made explicit within the analysis workflow. This makes it challenging for consumers of  the analysis to evaluate the results, since it becomes difficult to disentangle whether discrepancies arise from differing expectations or from the use of statistical technique, without running the analysis themselves. Non-expert analysts, lacking prior knowledge or instinct, may not have clear expectations of the results. This can lead to reduced confidence of the analysis and makes it more difficult and time-consuming to diagnose the cause of the deviation when the results don't align with expectations.

By explicitly formulating these expectations, an analysis plan can guide the analysis process, facilitate the communication and comparison with other analyses and evaluate the validity of the results.

* why it is useful to be explicit about expectation? 
    * and can be used to guide the analysis process
    * with simulation, the unit tests can be used to construct the "result universe" 
    * With formulated expectation, one can generate multiple version of the dataset for teaching statistics

When the expectation is formulated, we can also construct values fulfill or not fulfill the expectation. With simulation, this allows to construct the set of all possible results that can be obtained from the analysis process, which we called the result universe. With such a result universe, we can see how the results obtained by the analysts are situated in all possible results. When accompanied by the result universe, it is clearer how different factors in the analysis affects the main results. This can be seen as a direct towards trustworthy data analysis, where the audience of the analysis to exercise their own cognitive model [@grolemund_cognitive_2014] to evaluate the results reported.

>>>>

Experienced analysts often have implicit expectation about the outcome and rely on a few "directional signs" to check when the outcome deviate from those expectation. However, these expectations are rarely made explicit within the analysis workflow. This makes it challenging for consumers of  the analysis to evaluate the results, since it becomes difficult to disentangle whether discrepancies arise from differing expectations or from the use of statistical technique, without running the analysis themselves. Non-expert analysts, lacking prior knowledge or instinct, may not have clear expectations of the results. This can lead to reduced confidence of the analysis and makes it more difficult and time-consuming to diagnose the cause of the deviation when the results don't align with expectations. By explicitly formulating these expectations, an analysis plan can guide the analysis process, facilitate the communication and evaluate the validity of the results.

The expectations can be thought of as a set of unit tests used to validate the results of data analysis. By specifying a range of values for these tests, multiple versions of the dataset can be generated to satisfy different sets of plan expectations. This allows us to present what we called the "result universe" -- the complete set of possible results that can be obtained from one data analysis process. By visualizing the result universe, data analysis consumers can observe how changes in expectations affect the results and the range of alternative outcomes that could arise under different conditions. This enables them to evaluate the outcomes based on their own plan expectations and gain a broader perspective on how the actual results produced by analysts fit within this spectrum of possibilities, promoting transparency and trust in analysis.

Furthermore, by generating multiple versions of the data, we can emulate various scenarios within the same context for students to exercise judgement when conducting data analysis in a classroom setting.
